# Install the package.
```{r}
BiocManager::install("impute")
BiocManager::install("preprocessCore")
BiocManager::install("BioNERO")

install.packages("WGCNA")
```

# Load the package.
```{r}
library(WGCNA)
library(tidyverse)
library(ggpubr)
library(magrittr)
library(edgeR)
library(gtools)
library(data.table)
library(topGO)
library(Rgraphviz)
library(doParallel)
registerDoParallel(cores=8)
```

# Load the data, and trim away the columns that are not necessary. Eliminate as well the samples that we discarded based on the PCA.
```{r}
setwd("~/Desktop/PhD/Research/G.buenoi-DGE")
count_data <- read.table("./files/subread_counts_wingbuds.txt",header=TRUE, row.names = 1) # All genes that are expressed are included in the analysis.
#count_data <- read.table("./files/subread_counts_wingbuds_DE.txt",header=FALSE, row.names = 1) # Only DEG are included.
count_data <- count_data[,-1:-5]
colnames(count_data) <- c("A_4E_18_2", "A_4E_18_3", "A_4E_18_4", "A_4E_18_5", "A_4E_18_6", "A_4E_18_7", "B_4L_18_1", "B_4L_18_2", "B_4L_18_5", "B_4L_18_6", "B_4L_18_7", "C_5E_18_1", "C_5E_18_2", "C_5E_18_3", "C_5E_18_4", "C_5E_18_5", "C_5E_18_7", "D_5L_18_1", "D_5L_18_2", "D_5L_18_3", "D_5L_18_4", "D_5L_18_5", "D_5L_18_6", "W_5L_12_1", "W_5L_12_2", "W_5L_12_3", "W_5L_12_4", "W_5L_12_5", "X_4E_12_2", "X_4E_12_3", "X_4E_12_4", "X_4E_12_5", "X_4E_12_6", "X_4E_12_7", "Y_4L_12_1", "Y_4L_12_3", "Y_4L_12_5", "Y_4L_12_6", "Y_4L_12_7", "Y_4L_12_8", "Z_5E_12_1", "Z_5E_12_2", "Z_5E_12_3", "Z_5E_12_4", "Z_5E_12_5", "Z_5E_12_7")

count_data <- count_data[, -c(7,14,18,19,22,27,36)]
```

# Create the metadata matrix and the DGEList element. Filter for low expression and extract normalized counts.
```{r}
groups <- read.table("./files/groups_WGCNA.txt", header = TRUE, row.names = 1)
Group <- factor(paste(groups$Stage, groups$Photoperiod, sep = "_"))

dge_raw <- DGEList(counts = count_data, group = Group)

keep <- rowSums(cpm(dge_raw)>=5) >= 3
dge_filtered <- dge_raw[keep,,keep.lib.sizes=FALSE]

dge_filtered_TMM <- calcNormFactors(dge_filtered, method = "TMM")
#cpm(dge_filtered) # I calculated myself the cpm values after normalization and they are correct. Extracting the output of cpm() should suffice.

normalized_counts <- cpm(dge_filtered_TMM) # No log-transformation as recommended by the original author of WGCNA (https://peterlangfelder.com/2018/11/25/filtering-and-collapsing-data/)

rm(dge_filtered)
```

# Transpose the table to be able to use it as input in WGCNA.
```{r}
normalized_counts <- t(normalized_counts)
```

# Now this table should be enough to start running WGCNA.
```{r}
powers = c(c(1:10), seq(from = 12, to = 30, by = 2))

enableWGCNAThreads()
WGCNAnThreads()
enableWGCNAThreads(nThreads = 7)

sft <- pickSoftThreshold(
  normalized_counts,
  blockSize = 1000,
  powerVector = powers,
  dataIsExpr = TRUE,
  corFnc = cor,
  networkType = "signed",
  verbose = 10
)

rm(powers)

write.table(sft$fitIndices, "./WGCNA/soft_threshold.txt", quote=FALSE, sep = "\t", row.names = FALSE)
```

# We create a new variable for the signed R2, which is a measure of the fit of the model.
```{r}
sft_df <- data.frame(sft$fitIndices) %>%
  dplyr::mutate(model_fit = -sign(slope) * SFT.R.sq)

png("./WGCNA/soft_threshold.png")
plot(sft$fitIndices[, 1],
     -sign(sft$fitIndices[, 3]) * sft$fitIndices[, 2],
     xlab = "Soft Threshold (power)",
     ylab = "Scale Free Topology Model Fit, signed R^2",
     main = paste("Scale independence"),
     abline(h = 0.80, col = "red") # The threshold of 0.8 I found in https://alexslemonade.github.io/refinebio-examples/04-advanced-topics/network-analysis_rnaseq_01_wgcna.html#410_Which_modules_have_biggest_differences_across_treatment_groups
)
png("./WGCNA/soft_threshold.png")
dev.off()

rm(sft_df)

# We select the power = 16 for all genes as it is the first to exceed the recommended threshold of 0.8 signed R2. The signed R2 plateaus after that.
```

# Run the WGCNA.
```{r}
# For this chunk to work, it is necessary to restart R and load ONLY the WGCNA package.
bwnet <- blockwiseModules(normalized_counts,
  maxBlockSize = 5000, # What size chunks (how many genes) the calculations should be run in
  TOMType = "signed", # topological overlap matrix
  power = 16, # soft threshold for network construction, 16 for whole geneset, 12 for DE genes as well.
  numericLabels = TRUE, # Let's use numbers instead of colors for module labels
  randomSeed = 1234, # there's some randomness associated with this calculation
  # so we should set a seed
  minModuleSize = 30, # minimum number of genes per cluster.
  mergeCutHeight = 0.25, # clusters closer than 0.25 (equivalent to correlation of 0.75 or higher) will be merged
  verbose = 3
)
# Now load all packages again.
```
# We save the main WGCNA file to an object to be able to go back to it.
```{r}
readr::write_rds(bwnet,"./WGCNA/wgcna_results.RDS")
bwnet <- readRDS("./WGCNA/wgcna_results.RDS")
```

# We can plot the modules in a dendogram to see the proximity between different genes.
```{r}
# Convert labels to colors for plotting
mergedColors = labels2colors(bwnet$colors)
#mergedColors
# Plot the dendrogram and the module colors underneath
plotDendroAndColors(
  bwnet$dendrograms[[1]],
  mergedColors[bwnet$blockGenes[[1]]],
  "Module colors",
  dendroLabels = FALSE,
  hang = 0.03,
  addGuide = TRUE,
  guideHang = 0.05)

#bwnet$colors[bwnet$blockGenes[[1]]]
table(bwnet$colors) # How many genes cluster within each module.

module_eigengenes <- bwnet$MEs
module_eigengenes # This shows, for each WGCNA cluster of genes, the relative contribution to each sample (library).
sum(module_eigengenes$ME0) # They add to 0, as the contribution of each sample can be positive or negative (between 1 and -1)
```


# Load metadata.
```{r}
metadata <- read.table("./files/groups_WGCNA.txt",header=TRUE, row.names = 1)
metadata$Photoperiod <- as.factor(metadata$Photoperiod)
#all.equal(rownames(metadata), rownames(module_eigengenes)) # Check the names of the different rows (sample names) coincide.
```

# Enter the design matrix.
```{r}
des_mat <- model.matrix(~metadata$Stage*metadata$Photoperiod) # This replicates the matrix used in edgeR for DGE.
des_mat
colnames(des_mat) <- c("(Intercept)", "4L", "5E", "5L", "18", "4L:18", "5E:18", "5L:18")
```

# Output the cluster to which each gene belongs and save it in a file for future reference.
```{r}
module_df <- data.frame(
  gene_id = names(bwnet$colors),
  colors = labels2colors(bwnet$colors),
  module = bwnet$colors
)

write_delim(module_df,
            file = "./WGCNA/gene_modules.txt",
            delim = "\t")
```

```{r}
# lmFit() needs a transposed version of the matrix
fit <- limma::lmFit(t(module_eigengenes), design = des_mat)

# Apply empirical Bayes to smooth standard errors
fit <- limma::eBayes(fit)

# Apply multiple testing correction and obtain stats
stats_df <- limma::topTable(fit, number = ncol(module_eigengenes)) %>%
  tibble::rownames_to_column("module")

head(stats_df) # Here we can identify the modules that have more significant differences in gene expression across the different treatment comparisons.
write_delim(stats_df,
            file = "./WGCNA/stats_modules.txt",
            delim = "\t")
```



# Calculate eigengene (median gene value for each module) and plot them in a heatmap to see how they are associated with each treatment.
```{r}
# Get Module Eigengenes per cluster
MEs0_whole <- moduleEigengenes(normalized_counts, mergedColors, excludeGrey = TRUE)#$eigengenes
MEs0 <- moduleEigengenes(normalized_counts, mergedColors, excludeGrey = TRUE)$eigengenes
variation_explained <- moduleEigengenes(normalized_counts, mergedColors, excludeGrey = TRUE)$varExplained
# We see it is extremely low (0.13) for the grey cluster. Hence we can discard it. See "excludeGrey = TRUE" above)

# Reorder modules so similar modules are next to each other
MEs0 <- orderMEs(MEs0)
module_order = names(MEs0) %>% gsub("ME","", .)

# Add treatment names
MEs0$treatment = row.names(MEs0)

# tidy & plot data
mME = MEs0 %>%
  pivot_longer(-treatment) %>%
  mutate(
    name = gsub("ME", "", name),
    name = factor(name, levels = module_order)
  )

mME$name <- gsub('grey','0', mME$name)
mME$name <- gsub('turquoise','1', mME$name)
mME$name <- gsub('blue','2', mME$name)
mME$name <- gsub('brown','3', mME$name)
mME$name <- gsub('yellow','4', mME$name)
mME$name <- gsub('green','5', mME$name)
mME$name <- gsub('red','6', mME$name)
mME$name <- gsub('black','7', mME$name)
mME$name <- gsub('pink','8', mME$name)
mME$name <- gsub('magenta','9', mME$name)
mME$name <- gsub('purple','10', mME$name)
mME$name <- gsub('54','11', mME$name)
mME$name <- gsub('tan','12', mME$name)
mME$name <- gsub('salmon','13', mME$name)
mME$name <- gsub('cyan','14', mME$name)

mME$treatment <- gsub('A_4E','i4E', mME$treatment)
mME$treatment <- gsub('X_4E','i4E', mME$treatment)
mME$treatment <- gsub('B_4L','i4L', mME$treatment)
mME$treatment <- gsub('Y_4L','i4L', mME$treatment)
mME$treatment <- gsub('C_5E','i5E', mME$treatment)
mME$treatment <- gsub('Z_5E','i5E', mME$treatment)
mME$treatment <- gsub('D_5L','i5L', mME$treatment)
mME$treatment <- gsub('W_5L','i5L', mME$treatment)

# For display reasons only, create a new column with the "values" rounded to 2 decimals.
mME$rounded <-round(mME$value,2)
mME$name = factor(mME$name, levels = c("14", "13", "12", "11", "10", "9", "8", "7", "6", "5", "4", "3", "2", "1", "0"))
mME[with(mME, order("treatment")),]
mME$treatment = factor(mME$treatment, levels = c("i4E_12_2", "i4E_12_3", "i4E_12_4", "i4E_12_5", "i4E_12_6", "i4E_12_7", "i4L_12_1", "i4L_12_5", "i4L_12_6", "i4L_12_7", "i4L_12_8", "i5E_12_1", "i5E_12_2", "i5E_12_3", "i5E_12_4", "i5E_12_5", "i5E_12_7", "i5L_12_1", "i5L_12_2", "i5L_12_3", "i5L_12_5", "i4E_18_2", "i4E_18_3", "i4E_18_4", "i4E_18_5", "i4E_18_6", "i4E_18_7", "i4L_18_2", "i4L_18_5", "i4L_18_6", "i4L_18_7", "i5E_18_1", "i5E_18_2", "i5E_18_4", "i5E_18_5", "i5E_18_7", "i5L_18_3", "i5L_18_4", "i5L_18_6"))
mME[with(mME, order("treatment")),]
mME_mod <- with(mME, mME[order(treatment, name), ]) # We reorder the data based on treatment first, then name (number of the module). I manually checked this and the resulting plot with geom_tile is properly displayed :)

plot <- mME_mod %>% ggplot(., aes(x=treatment, y=name, fill=value)) +
  geom_tile() +
  theme_bw() +
  scale_fill_gradient2(
    low = "blue",
    high = "red",
    mid = "white",
    midpoint = 0,
    limit = c(-1,1)) +
  theme(plot.title=element_text(size = 15, face = "bold"),
        axis.text.x = element_text(angle=90, size=8, face = "bold"),
        axis.text.y=element_text(size=8, face = "bold"),
        axis.title = element_text(size = 12, face="bold"),
        legend.key.height = unit(0.5, "cm"),
        legend.title = element_blank(),
        legend.text = element_text(size=10)) +
  labs(y = "Module Number", x = "Library", fill="corr") +
  geom_text(aes(label = rounded), size = 2, angle=90) +
  guides(fill = guide_colorbar(barwidth = 0.5, barheight = 10))

plot

ggsave("./WGCNA/Module-trait.png", dpi = 1000)
ggsave("./WGCNA/Module-trait.pdf", dpi = 1000)
```


# Now we can explore specific modules that we find interesting. 
# Interesting clusters are somewhat arbitrarily determined, and might include those that have substantial differnece in the correlation of the eigengene in a given stage across photoperiods, or significant variation across time.
# Looking at the module-trait relationships we got, I would say the interesting are:
# Module 2, Module 5, Module 8 for being very different in i5L across photoperiods.
# Module 11, Module 14 for being different in i4L across photoperiods.
# Module 10 for being tightly positively correlated in i5E, where we don't see any DE.
```{r}
prefix_output="topGO-all"
module_number="14"



candidate_genes <- module_df %>% filter(module == module_number)
candidate_genes <- data.frame(candidate_genes$gene_id)
colnames(candidate_genes) <- "geneid"

geneNames_all <- read.table("geneNames_annotated_genes.tsv", header = F)
colnames(geneNames_all) <- "geneid"
geneNames_all <- as.character(geneNames_all$geneid)

geneList_all <- factor(as.integer(geneNames_all %in% candidate_genes$geneid))
table(geneList_all)
names(geneList_all) <- geneNames_all

geneID2GO_test <- readMappings(file="geneid2go_modified.map", sep = "\t")

topGOdata_BP_all <- new("topGOdata",
               ontology="BP",
               allGenes=geneList_all,
               annot=annFUN.gene2GO,
               gene2GO=geneID2GO_test,
               nodeSize=1)

topGOdata_CC_all <- new("topGOdata",
               ontology="CC",
               allGenes=geneList_all,
               annot=annFUN.gene2GO,
               gene2GO=geneID2GO_test,
               nodeSize=1)

topGOdata_MF_all <- new("topGOdata",
               ontology="MF",
               allGenes=geneList_all,
               annot=annFUN.gene2GO,
               gene2GO=geneID2GO_test,
               nodeSize=1)

resultFisher_BP_all <- runTest(topGOdata_BP_all, algorithm = "classic", statistic = "fisher")
resultKS_BP_all <- runTest(topGOdata_BP_all,algorithm="classic",statistic="ks")
resultKS.elim_BP_all <- runTest(topGOdata_BP_all,algorithm="elim",statistic="ks")

resultFisher_CC_all <- runTest(topGOdata_CC_all, algorithm = "classic", statistic = "fisher")
resultKS_CC_all <- runTest(topGOdata_CC_all,algorithm="classic",statistic="ks")
resultKS.elim_CC_all <- runTest(topGOdata_CC_all,algorithm="elim",statistic="ks")

resultFisher_MF_all <- runTest(topGOdata_MF_all, algorithm = "classic", statistic = "fisher")
resultKS_MF_all <- runTest(topGOdata_MF_all,algorithm="classic",statistic="ks")
resultKS.elim_MF_all <- runTest(topGOdata_MF_all,algorithm="elim",statistic="ks")

allRes_BP_all <- GenTable(topGOdata_BP_all, classicFisher = resultFisher_BP_all, classicKS = resultKS_BP_all, elimKS = resultKS.elim_BP_all, orderBy = "classicFisher", topNodes = length(score(resultFisher_BP_all)))

allRes_CC_all <- GenTable(topGOdata_CC_all, classicFisher = resultFisher_CC_all, classicKS = resultKS_CC_all, elimKS = resultKS.elim_CC_all, orderBy = "classicFisher", topNodes = length(score(resultFisher_CC_all)))

allRes_MF_all <- GenTable(topGOdata_MF_all, classicFisher = resultFisher_MF_all, classicKS = resultKS_MF_all, elimKS = resultKS.elim_MF_all, orderBy = "classicFisher", topNodes = length(score(resultFisher_MF_all)))

write.table(allRes_BP_all, file = c(paste0("./WGCNA/",prefix_output,"_BP_module_",module_number,"_results.txt")), sep = "\t")
write.table(allRes_CC_all, file = c(paste0("./WGCNA/",prefix_output,"_CC_module_",module_number,"_results.txt")), sep = "\t")
write.table(allRes_MF_all, file = c(paste0("./WGCNA/",prefix_output,"_MF_module_",module_number,"_results.txt")), sep = "\t")
```


#### PLOT FOR SPECIFIC MODULES ####

```{r}
#First step is to extract the name of the genes that belong to one cluster of interest.
gene_module <- read.table("./WGCNA/gene_modules.txt", header = TRUE)
genes_ME2 <- gene_module[gene_module$module == "8", ]
genes_ME2 <- genes_ME2$gene_id

# Now we filter the "normalized_counts" element (see above in the script), and retain ONLY the rows for the genes belonging to the module of interest.
cpm_ME2 <- data.frame(normalized_counts[rownames(normalized_counts) %in% genes_ME2,])

cpm_ME2 <- t(cpm_ME2)
i4E_18 <- cpm_ME2[1:6,]
i4L_18 <- cpm_ME2[7:10,]
i5E_18 <- cpm_ME2[11:15,]
i5L_18 <- cpm_ME2[16:18,]
i4E_12 <- cpm_ME2[23:28,]
i4L_12 <- cpm_ME2[29:33,]
i5E_12 <- cpm_ME2[34:39,]
i5L_12 <- cpm_ME2[19:22,]

i4E_18_means <- data.frame(colMeans(i4E_18))
i4L_18_means <- data.frame(colMeans(i4L_18))
i5E_18_means <- data.frame(colMeans(i5E_18))
i5L_18_means <- data.frame(colMeans(i5L_18))
i4E_12_means <- data.frame(colMeans(i4E_12))
i4L_12_means <- data.frame(colMeans(i4L_12))
i5E_12_means <- data.frame(colMeans(i5E_12))
i5L_12_means <- data.frame(colMeans(i5L_12))

means_18 <- cbind(i4E_18_means, i4L_18_means, i5E_18_means, i5L_18_means)
means_12 <- cbind(i4E_12_means, i4L_12_means, i5E_12_means, i5L_12_means)
colnames(means_18) <- c("i4E","i4L","i5E","i5L")
colnames(means_12) <- c("i4E","i4L","i5E","i5L")

means_18$Max <- pmax(means_18$i4E, means_18$i4L, means_18$i5E, means_18$i5L) # Find max of all time-points
means_12$Max <- pmax(means_12$i4E, means_12$i4L, means_12$i5E, means_12$i5L) # Find max of all time-points

means_18$i4E <- means_18$i4E / means_18$Max
means_18$i4L <- means_18$i4L / means_18$Max
means_18$i5E <- means_18$i5E / means_18$Max
means_18$i5L <- means_18$i5L / means_18$Max
means_12$i4E <- means_12$i4E / means_12$Max
means_12$i4L <- means_12$i4L / means_12$Max
means_12$i5E <- means_12$i5E / means_12$Max
means_12$i5L <- means_12$i5L / means_12$Max

means_18 <- data.frame(means_18[,1:4])
means_12 <- data.frame(means_12[,1:4])

means_18 <- t(means_18)
means_12 <- t(means_12)
groups <- c("i4E", "i4L", "i5E", "i5L")

means_18 <- data.table(cbind(groups, means_18))
#means_18 <- data.table(means_18[,1:100])
means_12 <- data.table(cbind(groups, means_12))
#means_12 <- data.table(means_12[,1:100])

means_18_melt <- melt.data.table(means_18, id.vars = "groups")
means_18_melt <- data.frame(means_18_melt)
means_18_melt = subset(means_18_melt, select = -c(variable))
means_18_melt$value <- as.numeric(means_18_melt$value)
means_18_melt$photoperiod <- as.factor("18L:6D")
means_12_melt <- melt.data.table(means_12, id.vars = "groups")
means_12_melt <- data.frame(means_12_melt)
means_12_melt = subset(means_12_melt, select = -c(variable))
means_12_melt$value <- as.numeric(means_12_melt$value)
means_12_melt$photoperiod <- as.factor("12L:12D")

means_melt <- rbind(means_12_melt, means_18_melt)

plot <- ggplot(means_melt, aes(x = groups, y = value, fill = photoperiod)) +
  theme_bw() +
  theme(panel.grid = element_blank(),
        axis.text.y = element_text(size=25),
        axis.text.x = element_text(size=25),
        axis.title = element_text(size = 25, face="bold"),
        legend.title = element_blank()) +
  geom_boxplot(notch=TRUE, outlier.alpha = 0.25) +
  #geom_point(position = position_jitterdodge(), alpha = 0.2, size = 0.25) + # Add points for each gene.
  scale_fill_manual(values = c("lightblue", "orange")) +
  xlab("Developmental stage") + ylab("Fraction max")

plot

ggsave("./WGCNA/ME8.pdf", plot = plot, dpi = 1000, height = 5, width = 7)
ggsave("./WGCNA/ME8.png", plot = plot, dpi = 1000, height = 5, width = 7)
```


